---
title: "Spotify"
author: "Mason Ogden"
date: "February 1, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(stringr)
spot <- read.csv('C:/Users/mogde/Documents/spotify_playlist.csv')
spot_shuffled <- read.csv('C:/Users/mogde/Documents/spotify_shuffled.csv')
names(spot) <- c('song_name','artist','artist_code','album_name','album_num')
```
I used one of my own playlists 'Guitars and Keyboards' for this analysis. My playlist has 18 artists with 73 songs from a total of 34 albums. To get my observed shuffled playlist, I clicked 'SHUFFLE PLAY' and then recorded the order in which the songs played, as well as the artist and the album. 

Using the sample() function, I will randomly shuffle the ordered spotify playlist 1000 times and look
for runs where the same artist was played 3 or more times in a row. 
```{r}
nreps <- 1000
thousand_shuffles <- lapply(1:nreps, function(x) as.character(sample(spot$artist_code)))
# 'thousand_shuffles' is a list with 1000 entries; every entry being a list with 73 shuffled letters in it. 
thousand_run_lengths <- rep(list(''), nreps) # Initialize list
for (i in 1:nreps) { # For each of the 1000 shuffled playlists in thousand_shuffles
  thousand_run_lengths[[i]] <- rle(thousand_shuffles[[i]])$lengths # Get the lengths of the longest run of songs for every artist
}
observed_run_length <- rle(as.character(spot_shuffled$artist_code))$lengths  # Get run lengths for the playlist shuffled by spotify
obs_num_3_or_more <- sum(observed_run_length >= 3) # Count how many of those spotify run lengths were 3 songs or longer
sim_num_3_or_more <- rep(list(''), nreps) # Initialize list
for (j in 1:nreps) { # For each of the 1000 vectors of run lengths
  sim_num_3_or_more[[j]] <- sum(thousand_run_lengths[[j]] >= 3) # Calculate how many runs were 3 songs or longer for each shuffle
}
sim_num_3_or_more <- unlist(sim_num_3_or_more) # Turn the list of runs into one long vector
mean_sim_run_length <- mean(sim_num_3_or_more)# Find the mean length  of all simulated runs
```


```{r}
sim_3_data <- data.frame(value=sim_num_3_or_more)
names(sim_3_data) <- 'num_3_or_more'
ggplot(sim_3_data, aes(x=num_3_or_more)) + 
  geom_histogram(binwidth=1, fill='coral') + 
  xlab('Number of Runs of 3 or More') +
  ylab('Count') + 
  ggtitle('Distribution of Number of Runs of 3 Songs or More \nby the Same Artist in 1000 Shuffled Playlists')

```

So, I know that there were 3 runs of 3 or more in my playlist that was shuffled by Spotify. What is the probability of getting this many runs of at least 3 if the playlist order was \textit{really} random?
```{r}
sim_p_value <- sum(sim_num_3_or_more >= obs_num_3_or_more) / nreps
sim_p_value
```

The probability that getting at least 3 songs by the same artist 3 separate times if the shuffling method were truly random is 0.018. Based on this simulated p-value, I conclude that clicking 'Shuffle Play' on a playlist on Spotify does not truly randomize the order of songs. 